From Freebase to Wikidata: The Great Migration

### 2015-10-08 paper.tex
Reworked Section 3.4.1.

The data quality of Freebase was examined thoroughly by the Wikidata community
and anecdotal evidence indeed confirmed that there were certain challenges.
For example, a~small city in France was assigned an ISO country code,%
\footnote{See \url{http://www.freebase.com/m/03cc0d0}}
then Boston (the city in Massachusetts) had the type \texttt{/people/person},
and several other occurrences of data quality issues were revealed.
In consequence, a~fully automatic upload of all the contents from Freebase into Wikidata
did not seem advisable as the expectations of the Wikidata community regarding the
quality of automatically uploaded data are high (and rightly so).
Despite the fact that an automated upload of all the data
would have led to accelerated growth in the short term,
such a~fast increase of available data would certainly have endangered
the sense of ownership of the Wikidata community for its data.
As an alternative approach, we decided to rely on crowdsourced human curation
and created the \emph{Primary Sources Tool},
a~software solution that displays Freebase statements
for verification to the user that can be added to the currently shown Wikidata item.
The tool will be described in more detail in \autoref{sec:primary-sources-tool},
we provide a~short summary of its functionality here.
\autoref{fig:primary-sources-tool} shows a~screenshot of the user interface
of the tool that integrates nicely into the Wikidata user interface.
The user can, with just one click, reject or approve a~statement,
and, in case of approval, therewith add the statement to Wikidata.
The Primary Sources Tool is currently available as a~so-called gadget,
\emph{i.e.}, as an opt-in feature in Wikidata.
This tool has no dependencies on Freebase and can be easily reused for other projects
that want to import datasets into Wikidata.
Therefore, we have implemented an upload API in the back-end
and a~filter in the front-end that allows users to display only statements from a~given dataset.
This flexibility is much-wanted, and during the development period of the project
another researcher approached us in order to upload a~dataset extracted from natural language text.
To visualize the migration progress, we have created a~realtime dashboard
available at \url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}.
The code of the Primary Sources Tool is openly available
(\url{https://github.com/google/primarysources}) under the terms of the Apache~2.0 license.
At the time of writing (October, 2015), the tool has been used by more than hundred users
who performed about 30,000 approval or rejection actions.
More than 12M~statements have been uploaded in total.

\subsection{Mappings}

@@

Another challenge is to get a~good mapping between Freebase topics and properties
and Wikidata items and properties.
Two mappings between Freebase topics and Wikidata items were initially available,
and we worked on further mappings.
A first one has been created by Google in October 2013%
\footnote{Freebase mappings \url{http://developers.google.com/freebase/data}}
and is based on Wikipedia links already present in Freebase: if a~Freebase topic and
a Wikidata items share at least two Wikipedia links, then they are about the same subject.
The quality of this mapping is considered as very good by the Wikidata community
who have uploaded it to the wiki and further curated and maintained it by hand.
It maps 1.15M~of Freebase topic.
A second mapping has been created by Samsung, and is actively maintained%
\footnote{Samsung mappings \url{http://github.com/Samsung/KnowledgeSharingPlatform}.}.
It is based on the same idea but matches a~Freebase topic with a~Wikidata item
even if there is only a~single Wikipedia link shared.
The quality of the mapping is lower (there are some wrong matches,
often because there is no clear separation between topics and disambiguation pages in Wikipedia)
but it provides 4.4M~links.
As we have seen before, our quality target is not very high,
so we have chosen to merge the two dumps and, for the 6,000 conflicts,
to prioritize the mapping given by Wikidata.
To improve the resulting mapping we have also tried to do some reconciliation
based on external database IDs shared by Freebase and Wikidata, like MusicBrainz, VIAF, \emph{etc.}
With this technique we were able to match 0.8M~IDs,
creating a~mapping for 0.6M~topics---most of them already in our dataset.
This resulted in an additional 84,000 items mapped.
We also used data from Google's Knowledge Graph to add 0.1M~further mappings
matching also a~topic and an item if they share a~Wikipedia link.
We could maybe add some more data by looking at functional relations like ``father'' and ``mother''.
But these relations only apply to relatively small sets of data (like well known families)
and the functional assumption may run into some edge cases.
For example, the \texttt{/people/person/parents} property in Freebase also covers also stepparents,
creating cases where a~person has more than one male or female \texttt{/people/person/parents}.
Eventually, with these four sources we mapped 4.56M~items.
We believe that by doing so we have mapped most of the topics that could be automatically mapped
as suggested by the small increases added by the ID reconciliation approach
and the Knowledge Graph data.
Furthermore, the mapped topics have an average number of facts of about~13.9,
whereas the topics that were not mapped have an average of 5.7~facts.
We have mapped the majority of topics with more than 80~facts.
This allows us to conclude that we have mapped the most important items of Freebase.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/facts-topics-mapping.pdf}
\caption{Number of mapped and not mapped topics with more than x facts.
Note the log scale on both axis.}
%Data sheet: https://docs.google.com/spreadsheets/d/1jYGVIqAmyouHKjvhM6FbpgTD8bb5tkJy9PEMELkuMwo/edit
\end{figure}

@@

>>>>

The data quality of Freebase was examined thoroughly by the Wikidata community
and anecdotal evidence indeed confirmed that there were certain challenges.
For example, a~small city in France was assigned an ISO country code,%
\footnote{Freebase item page of the city of Saint-Martin: \url{http://www.freebase.com/m/03cc0d0}}
then Boston (the city in Massachusetts) had the type \texttt{/people/person},
and several other occurrences of data quality issues were revealed.
In consequence, a~fully automatic upload of all the contents from Freebase into Wikidata
did not seem advisable as the expectations of the Wikidata community regarding the
quality of automatically uploaded data are high (and rightly so).
Despite the fact that an automated upload of all the data
would have led to accelerated growth in the short term,
such a~fast increase of available data would certainly have endangered
the sense of ownership of the Wikidata community for its data.
As an alternative approach, we decided to rely on crowdsourced human curation
and created the \emph{Primary Sources Tool},
a~software solution that displays Freebase statements
for verification to the user that can be added to the currently shown Wikidata item.
The tool will be described in more detail in \autoref{sec:primary-sources-tool}.

\subsection{Data Mappings}

@@

The last challenge concerns the data mappings between Freebase topics and properties
and Wikidata items and properties.
Two mappings between Freebase topics and Wikidata items were initially available,
and in addition to those, we worked on further mappings.
A~first one has been created by Google in October 2013%
\footnote{Freebase mappings: \url{http://developers.google.com/freebase/data}}
and is based on Wikipedia links already present in Freebase: if a~Freebase topic and
a~Wikidata item share at least two Wikipedia links, they are assumed to describe the same subject.
The quality of this mapping is considered very good by the Wikidata community,
who thus have uploaded it to the Wiki and further curated and maintained it by hand.
It maps 1.15M~Freebase topics to their corresponding Wikidata items.
A~second actively maintained mapping has been created by Samsung.%
\footnote{Samsung mappings: \url{http://github.com/Samsung/KnowledgeSharingPlatform}}
It is based on the same idea, but matches a~Freebase topic with a~Wikidata item
even if there is only a~single shared Wikipedia link.
The precision of the mapping is naturally lower compared to Google's mapping,
as there are more wrong matches (often because there is no clear separation between topics
and disambiguation pages in Wikipedia), however, the recall increases to 4.4M~links.
As we have seen before \todo{Where?}, our quality target is not very high \todo{We should never write this},
so we have chosen to merge the two dumps \todo{The mappings dumps I suppose?} and, for the resulting 6,000 conflicts,
decided to prioritize the mapping given by Wikidata \todo{That is, Google's?}.
To improve the resulting mapping, we have also tried to do some reconciliation
based on external database IDs shared by Freebase and Wikidata, like MusicBrainz \todo{Citation needed}, VIAF \todo{Citation needed}, \emph{etc.}
With this technique, we were able to match 800K~IDs \todo{IDs? Is that MusicBrainz or VIAF IDs?},
creating a~mapping for 600K~topics---most of them already in our dataset \todo{Which dataset?}.
This resulted in an additional 84K items mapped.
We also used data \todo{What data exactly?} from Google's Knowledge Graph \todo{Citation desirable} to add 100K~further mappings
by also matching a~topic and an item if they share a~Wikipedia link \todo{Wasn't it leveraging the Knowledge Graph?}.
There is some potential to add some more data by looking at functional relations
like ``father'' and ``mother'', but these relations only apply to relatively small sets of data
(like well known families) and the functional assumptions may run into some edge cases.
For example, the \texttt{/people/person/parents} property in Freebase also covers stepparents,
which may create cases where a~person has more than one male or female \texttt{/people/person/parents}.
Eventually, with these four sources \todo{Google, Samsung, MusicBrainz, VIAF, Knowledge Graph. 5 Sources?}, we mapped 4.56M~items.
To the best of our knowledge, therewith we have already mapped
most of the topics that could be mapped automatically, 
which is also backed by the comparably small increases gained from the ID reconciliation approach \todo{Referring to MusicBrainz and VIAF?}
and the Knowledge Graph data.
Furthermore, the mapped topics have an average number of facts of about~13.9,
whereas the topics that were not mapped have an average number of 5.7~facts.
We have mapped the majority of topics with more than 80~facts.
This allows us to conclude that we have mapped the most important items of Freebase \todo{Easy "reviewer 2" question: why 80?}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/facts-topics-mapping.pdf}
\caption{Number of mapped and not mapped topics with more than x facts.
Note the log scale on both axes.}
%Data sheet: https://docs.google.com/spreadsheets/d/1jYGVIqAmyouHKjvhM6FbpgTD8bb5tkJy9PEMELkuMwo/edit
\end{figure}

@@

As outlined earlier, the \emph{Primary Sources Tool} is a~crowdsourced human curation
software solution that displays Freebase statements
for verification to the user that can be added to the currently shown Wikidata item.
The user can, with just one click, reject or approve a~statement,
and, in case of approval, therewith add the statement to Wikidata.
The Primary Sources Tool is currently available as a~so-called gadget,
\emph{i.e.}, as an opt-in feature in Wikidata.
This tool has no dependencies on Freebase and can be easily reused for other projects
that want to import datasets into Wikidata.
Therefore, we have implemented an upload API in the back-end
and a~filter in the front-end that allows users to display only statements from a~given dataset.
This flexibility is much-wanted, and during the development period of the project
another researcher approached us in order to upload a~dataset extracted from natural language text.
To visualize the migration progress, we have created a~realtime dashboard
available at \url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}.
The code of the Primary Sources Tool is openly available
(\url{https://github.com/google/primarysources}) under the terms of the Apache~2.0 license.
At the time of writing (October, 2015), the tool has been used by more than hundred users
who performed about 30,000 approval or rejection actions.
More than 12M~statements have been uploaded in total.
\autoref{fig:primary-sources-tool} shows a~screenshot of the user interface
of the tool that integrates nicely into the Wikidata user interface.
